---
title: "527_Final_Project"
author: "Andrew Peters"
date: "2024-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggplot2)
```
- Importing csv into tidyverse dataframes. For future reference, It would be a lot easier to normalize naming of columns and df's right off the bat.
```{r }
pokedex <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_data.csv")
gym_moves <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\gym_pkmn_moves.csv")
main_pkmn <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\main_pkmn_dataset_Apr30.csv")
effectiveness <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_effectiveness.csv")
pokemon_exp <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_ev_exp.csv")
evolution <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_evolution.csv")
trainers <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_platinum_trainers.csv")
pokemon_summary <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\pokemon_summary.csv")
locations_gym_sections <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\locations_gym_sections.csv")
exp_types <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\exp_types.csv")
trainers_done <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\trainers_done.csv")
pokemon_platinum <- read.csv("C:/Users/abpet/Desktop/STAT 527 Group Project\\Pokemon_Platinum_Names.csv")


pokedex_df <- data.frame(pokedex) # pokedex
gym_moves_df <- data.frame(gym_moves) # Moves used in gyms
main_pkmn_df <- data.frame(main_pkmn) # Main pokemon DF with all info USE THIS
effectiveness_df <- data.frame(effectiveness) #Type effectivness chart
pokemon_exp_df <- data.frame(pokemon_exp) #Base exp of pokemon needed for EXP gain formula
evolution_df <- data.frame(evolution) #Criteria for when pokemon evolve
trainers_df <- data.frame(trainers) # All trainers and their pokemon in plat. needs cleaning and gym sections.
pokemon_summary_df <- data.frame(pokemon_summary) # Large set of what looks like pokedex
locations_gym_sections_df <- data.frame(locations_gym_sections) #gym section 0 => obtainable before first gym
exp_types_df <- data.frame(exp_types) # Chart for exp groups, connect main_pkmn to give individuals their groups.
trainers_done_df <- data.frame(trainers_done) #Finalized platinum trainer info, pokemon lvl, gym section.
pokemon_platinum_df <- data.frame(pokemon_platinum)

main_pkmn_df; head(locations_gym_sections_df); head(evolution_df); head(pokemon_exp); head(pokemon_platinum_df); head(trainers_done_df); colnames(trainers_done_df); rownames(exp_types_df)<-exp_types_df$Level; exp_types_df; colnames(main_pkmn_df); exp_types_df;
```
Need to clean up main_pkmn_df, there are 600 unneeded rows.
```{r }
main_pkmn_df <- main_pkmn_df %>% filter(!is.na(hp)) #should filter out all 600+ unneeded rows
```

This was used to set the gym section for trainers to set gym section
```{r }
get_gym_section <- function(pokemon_level) {
  gym_leader_ace <- c(14, 22, 26, 32, 37, 41, 44, 50)
  for (i in seq_along(gym_leader_ace)) {
    if (pokemon_level < (gym_leader_ace[i] - 4)) {
      return(i-1)
    }
   
  }
  return(length(gym_leader_ace))  # Return the last section if level is high
}

```

This and most of the other work was done to clean and organize the trainer dataset.
```{r }
extract_levels <- function(pokemon_string) {
  # This will return NA where pokemon_string is empty and extract digits where it's not
  ifelse(nzchar(pokemon_string), sub(".*Level (\\d+)$", "\\1", pokemon_string), NA)
}


df <- trainers_df %>%
  mutate(across(starts_with("Pokemon"), extract_levels, .names = "level_{.col}")) %>%
  rowwise() 

df1 <- df %>%
  mutate(across(starts_with("level_Pokemon"), as.numeric)) %>% # Ensure numeric
  rowwise() %>%
  mutate(
    Max.Level = max(c(level_Pokemon1, level_Pokemon2, level_Pokemon3, level_Pokemon4, level_Pokemon5, level_Pokemon6), na.rm = TRUE),
    Gym.Section = get_gym_section(Max.Level)
  )

write.csv(df1,'trainers_done.csv')

```


Looking at availability of EXP throughout Pokemon Platinum.
Pokemon EXP groups formulas are below. Where EXP groups determine exp gain and required exp to reach level 100.
N = level and the functions return needed exp to level up. This isn't needed as I have exp_types df which tells how much exp is needed for each level.

```{r }
EXP_gain <- function(b,L,s=1,e=1,a=1.5,t=1) {
  # b = base exp yield
  # L = Level of fainted pokemon
  # s = number of pokemon participating in the battle
  # e = 1.5 if lucky egg else 1, we will assume no lucky egg
  # a = 1 if fighting wild pokemon, else 1.5. For this we will always assume trainer battle
  # t = 1.5 if traded pokemon, will assume no trades allowed.
  return(round(((b*L)/7) * (1/s) * e *a * t))
}
EXP_gain(54,6)#Test
EXP_gain(56,7)#Test
```
Tried to clean up the trainers df and spread it out to more columns. It started out as all the pokemon and levels in one column. Ended with pokemon names and levels all separated.
```{r }
trainers_done_df <- trainers_done_df %>%
  mutate(across(starts_with("Pokemon"), ~str_extract(.x, "^[^ ]+")))
head(trainers_done_df)
```

Had a lot of trouble with this section, getting things to work how I wanted. I initially tried to just have a single column for total EXP but the calculation was not working correctly. It would ignore the first pokemon column entirely. And it would also miscalculate the EXP gain. I have no idea why. So to simplify things I just calculated the individual EXP gain from each trainer pokemon and I can then easily sum the individual EXP for each trainer pokemon to have a total exp per trainer.
```{r }
names(trainers_done_df) <- tolower(gsub(" ", "", names(trainers_done_df)))
names(pokemon_exp_df) <- tolower(gsub(" ", "", names(pokemon_exp_df)))

trainers_exp_df <- trainers_done_df %>%
  rowwise() %>%
  mutate(
    exp_pokemon1 = {
      pokemon_name <- pokemon1
      pokemon_level <- as.numeric(level_pokemon1)
      if (!is.na(pokemon_name) && pokemon_name != "") { 
        base_exp <- pokemon_exp_df %>%    # Goes into pokemon_exp_df to pull base exp for exp_gain function later
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    exp_pokemon2 = {
      pokemon_name <- pokemon2
      pokemon_level <- as.numeric(level_pokemon2)
       if (!is.na(pokemon_name) && pokemon_name != "") {
        base_exp <- pokemon_exp_df %>%
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    exp_pokemon3 = {
      pokemon_name <- pokemon3
      pokemon_level <- as.numeric(level_pokemon3)
       if (!is.na(pokemon_name) && pokemon_name != "") {
        base_exp <- pokemon_exp_df %>%
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    exp_pokemon4 = {
      pokemon_name <- pokemon4
      pokemon_level <- as.numeric(level_pokemon4)
       if (!is.na(pokemon_name) && pokemon_name != "") {
        base_exp <- pokemon_exp_df %>%
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    exp_pokemon5 = {
      pokemon_name <- pokemon5
      pokemon_level <- as.numeric(level_pokemon5)
       if (!is.na(pokemon_name) && pokemon_name != "") {
        base_exp <- pokemon_exp_df %>%
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    exp_pokemon6 = {
      pokemon_name <- pokemon6
      pokemon_level <- as.numeric(level_pokemon6)
       if (!is.na(pokemon_name) && pokemon_name != "") {
        base_exp <- pokemon_exp_df %>%
                    filter(tolower(name) == tolower(pokemon_name)) %>%
                    pull(exp)
        if (length(base_exp) > 0 && !is.na(pokemon_level)) {
          EXP_gain(base_exp, pokemon_level)
        } else {
          0
        }
      } else {
        0
      }
    },
    total_EXP = exp_pokemon1 + exp_pokemon2 + exp_pokemon3 + exp_pokemon4 + exp_pokemon5 + exp_pokemon6
  ) %>%
  ungroup()


trainers_exp_df <- trainers_exp_df %>%
  select(trainer.name, number.of.pokemon, starts_with("exp_"), everything())


trainers_exp_df
write.csv(trainers_exp_df,'trainers_done_hopefully.csv')
```

Now I can finally start some analysis of EXP availability. I can sum each gym section, and visualize EXP progression throughout the game. Chose to use plotly so we can 
```{R }
# Summarize total EXP by gym section
exp_by_gym <- trainers_exp_df %>%
  group_by(gym.section) %>%
  summarise(total_EXP = sum(total_EXP, na.rm = TRUE)) %>%
  ungroup()


plot <- plot_ly(exp_by_gym, x = ~gym.section, y = ~total_EXP, type = 'bar',
                marker = list(
                              line = list(color = 'rgba(50, 171, 96, 1.0)', width = 1)),
                text = ~total_EXP, 
                textposition = 'outside') %>%
        layout(title = "Total EXP Available by Gym Section",
               xaxis = list(title = "Gym Section"),
               yaxis = list(title = "Total EXP"),
               hovermode = "closest")


plot
```
Now to look at the EXP gain of pokemon.
Just realized I made this more difficult than I needed to. The left half of this data set is the required EXP to reach the next level. Instead I took the EXP needed to reach, for example, 15 and subtracted the level I was at, say 12.
After writing the sentence above, I think I changed my mind again and like the way I did it.
```{r }
#Changes first row to be the new column names. as.character may have been unnecessary,can't remember, need to comment better.
colnames(exp_types_df) <- as.character(unlist(exp_types_df[1, ]))

# Remove the first row
exp_types_df <- exp_types_df[-1, ]
exp_types_df
```
I want pokemon summary to have the actual group name for exp gain, not just he number needed to reach lvl 100
This was good because I added it to the main_pkmn
```{R }
specific_rows <- exp_types_df[100,]

# Convert these rows to a long format for easier lookup
exp_types_long <- specific_rows %>%
  pivot_longer(
    cols = everything(), 
    names_to = "column_name", 
    values_to = "growth_value"
  ) %>%
  mutate(growth_value = as.integer(growth_value)) %>%  # Convert growth values to integer
  filter(!is.na(growth_value), column_name != "Level")

exp_types_long <- exp_types_long %>%
  group_by(column_name) %>%
  filter(growth_value == max(growth_value)) %>%
  ungroup()
exp_types_long

#join on growth value, looks for EXP needed to reach 100 in pkmn_sum and finds the growth_type in exp_types_long
pokemon_summary_df <- pokemon_summary_df %>%
  left_join(exp_types_long, by = c("experience_growth" = "growth_value"))

#Write to csv but then I added it to Ethan's main pkmn csv for easy access for everyone.
pokemon_summary_df
write.csv(pokemon_summary_df,'pokemon_summary_growth_type.csv')
```

Looking at availability of pokemon of different growth types throughout the game. Decided to do pokemon exclusive to gym sections, and cumulatively. In the end I decided to do cumulative because that is generally how people play the game.
```{r }

# Exclusive Dataframe
get_exclusive_section_data <- function(section) {
  main_pkmn_df %>% 
    filter(Gym.Section == section)
}

# Cumulative Dataframe
get_cumulative_section_data <- function(section) {
  main_pkmn_df %>%
    filter(Gym.Section <= section)
}

# Define gym leader ace levels
gym_leader_ace <- c(14, 22, 26, 32, 37, 41, 44, 50)

# Prepare dataframes for each section
exclusive_data <- lapply(0:max(main_pkmn_df$Gym.Section), get_exclusive_section_data)
cumulative_data <- lapply(0:max(main_pkmn_df$Gym.Section), get_cumulative_section_data)

calculate_growth_type_numbers <- function(df, ace_level) {
  df %>%
    #filter(Average.Level <= ace_level) %>% # I may  want to change this
    group_by(growth_type) %>%
    summarise(count = n())  
}

#These next two lines probably could have been done differently since I removed filter above.
# Apply analysis for exclusive data
exclusive <- mapply(calculate_growth_type_numbers, exclusive_data, gym_leader_ace, SIMPLIFY = FALSE)

# Apply analysis for cumulative data
cumulative <- mapply(calculate_growth_type_numbers, cumulative_data, gym_leader_ace, SIMPLIFY = FALSE)

# Convert to dataframe
exclusive_df <- bind_rows(exclusive, .id = "Gym.Section")
cumulative_df <- bind_rows(cumulative, .id = "Gym.Section")

# Pivoting for a better format for visualization
exclusive_long <- exclusive_df %>%
  pivot_longer(-c(Gym.Section, growth_type), names_to = "metric", values_to = "value")

cumulative_long <- cumulative_df %>%
  pivot_longer(-c(Gym.Section, growth_type), names_to = "metric", values_to = "value")

ggplot(exclusive_long, aes(x = Gym.Section, y = value, fill = growth_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Exclusive Gym Section Pokemon Availability",
       x = "Gym Section",
       y = "Number of Pokémon")

ggplot(cumulative_long, aes(x = Gym.Section, y = value, fill = growth_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Cumulative Gym Section Pokemon Availability",
       x = "Gym Section",
       y = "Number of Pokémon")

```
Now we can look at how many pokemon we can level up to the gym leader ace level in each gym section.
Using exclusive_df which is the number of pokemon of each growth type each gym section. Such that, pokemon exclusively caught in that gym section is considered. Later we can do cumulative.

From the dfs below we can find how many pokemon of each growth type are able to be leveled to the gym_leader_ace.
We will assume the starting level for each gym section is the leader ace before it (i.e starting level for gym4 is gymleader 3 ace.)
```{r }
#Put these here so I could look at the easier.
trainers_exp_df # Can find total EXP for each gym section

#exclusive_df # Number of growth type pokemon found in each gym section.

cumulative_df # Cumulative number of growth types

exp_by_gym

exp_types_df <- exp_types_df[1:6] %>% mutate_all(as.integer)

starting_level <- c(5, gym_leader_ace[1:7])


#returns a matrix of exp needed to reach the next level cap
exp_needed <- sapply(1:length(gym_leader_ace), function(i) {  
  unlist(exp_types_df[gym_leader_ace[i], ] - exp_types_df[starting_level[i], ])
})
exp_needed_df <- data.frame(exp_needed)
colnames(exp_needed_df) <- as.numeric(c(0:7))

# This df gives the amount of EXP needed to get from the gym_1 ace level to gym_2 ace level...
exp_needed_df

exp_analysis <- exp_needed_df %>%
  pivot_longer(
    cols = everything(),  # This ensures all columns are included
    names_to = "gym.section", 
    values_to = "EXP_Needed"
  ) %>%
  mutate(gym.section = as.integer(gsub("\\D", "", gym.section))) 


exp_analysis <- exp_analysis %>%
  left_join(exp_by_gym, by = "gym.section")
exp_pokemon_can_level <- exp_analysis %>% mutate(pokemon_can_level=total_EXP/EXP_Needed)
exp_needed_df
exp_analysis

exp_pokemon_can_level <- exp_pokemon_can_level %>% mutate(growth_type=rep(rownames(exp_needed_df),each=8, length.out=n()))




#This looks at individual leveling groups
ggplot(exp_pokemon_can_level, aes(x = gym.section, y = pokemon_can_level, fill = growth_type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Pokemon Leveling Capability by Growth Type and Gym Section",
       x = "Gym Section",
       y = "Number of Pokemon that Can Reach Gym Leader Ace Level",
       fill = "Growth Type") +
  facet_wrap(~growth_type, scales = "free_y")  


#Average number of pokemon that can reach the next gym leader ace level. Doesn't really consider growth_type
average_levels <- exp_pokemon_can_level %>%
  group_by(gym.section) %>%
  summarise(average_pokemon_can_level = mean(pokemon_can_level, na.rm = TRUE))
average_levels

#This looks at the average number of pokemon you can level up to gym leader ace.
ggplot(average_levels, aes(x = as.factor(gym.section), y = average_pokemon_can_level, fill = as.factor(gym.section))) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(title = "Average Number of Pokémon That Can Reach Gym Level by Gym Section",
       x = "Gym Section",
       y = "Average Pokémon Can Level") +
  scale_fill_discrete(name = "Gym Section")  
```












