---
title: "STAT 527 Final Project Work_AS"
author: "Anna Szlembarski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```{r}

gym_moves <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/gym_pkmn_moves.csv")
pokemon_types <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/main_pkmn_dataset.csv")
pokemon <- read.csv("C:/Users/szlem/Desktop/STAT 527/pokemon_data_copy.csv")
effectiveness <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/pokemon_effectiveness.csv")
locations <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/Updated_Locations_with_Gym_Sections.csv")
platinum_pokemon <- read.csv("C:/Users/szlem/Desktop/STAT 527/Pokemon_Platinum_Names_with_Type.csv")
trainers <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/Pokemon_Platinum_Trainers.csv")
experience <- read.csv("C:/Users/szlem/Desktop/STAT 527/STAT_527_Final_Project/csv/Pokemon_EV_EXP.csv")

library(ggplot2)
library(dplyr)
library(tidyverse)

rock_weakness <- c("Water", "Grass", "Fighting", "Ground", "Steel")

advantageous_pokemon <- platinum_pokemon %>%
  filter(Type1 %in% rock_weakness | Type2 %in% rock_weakness)

advantageous_pokemon_stats <- merge(advantageous_pokemon, pokemon, by = "Name", all.x = TRUE)

best_team <- advantageous_pokemon_stats %>%
  group_by(Name) %>%
  summarize(Total_Stats = sum(HP, Attack, Defense, SpAtk, SpDef, Speed)) %>%
  top_n(6, Total_Stats) %>%
  arrange(desc(Total_Stats)) %>%
  pull(Name)

print(best_team)
#this is without the XP or locations factored in, so I will need to add that in
```

Data Visualization
```{r}
# creating a chart to show types v base stats

platinum_pokemon_stats$AverageStats <- rowMeans(platinum_pokemon_stats[, c("HP", "Attack", "Defense", "SpAtk", "SpDef", "Speed")])

# combine Type1 and Type2 into one column
types <- unique(c(platinum_pokemon_stats$Type1, platinum_pokemon_stats$Type2))

# create a new data frame to store the types for each Pokemon
type_data <- platinum_pokemon_stats %>%
  select(Name, Type1, Type2, AverageStats) %>%
  pivot_longer(cols = c(Type1, Type2), names_to = "Type") %>%
  filter(!is.na(value)) %>%
  select(-Name)

# average stats for each type
type_avg_stats <- type_data %>%
  group_by(value) %>%
  summarise(AverageStat = mean(AverageStats, na.rm = TRUE))

ggplot(type_avg_stats, aes(x = reorder(value, AverageStat), y = AverageStat, fill = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Type", y = "Average of Base Stats", title = "Average of Base Stats for Each Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


